Studying the Performance of the Jellyfish Optimiser for the Application of Projection Pursuit

### 2024-06-03 optim.qmd
finalise the calculation for example 2

```{r setup, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(ggh4x)
```

@@

@@

```{r}
#| label: tbl-smoothness-squintability
#| tbl-cap: Parameters estimated from the Gaussian process (including variance, range, smoothness, and nugget) and scaled logistic function (theta1 - theta4) for both the pipe-finding and sine-wave finding problems. The "smooth" and "theta3" column represent the smoothness and squintability measures. 
load(here::here("data/smoothness.rda"))
load(here::here("data/squintability.rda"))
library(kableExtra)
tibble(shape = c(rep("pipe", 4), rep("sine", 8)), smoothness) |>
  left_join(squintability) |> 
  rename(d = n, smooth = smoothness) |> 
  knitr::kable(digits = 3) |> 
  column_spec(1, border_left = TRUE) |> 
  column_spec(c(3, 7, 11), border_right = TRUE) |> 
  kable_styling(latex_options = "hold_position")
```

@@

```{r eval = FALSE}
#| label: tbl-all-sim
#| tbl-cap: The first five rows of the datasets after combining all the calculation. The smoothness and squintability variable are uniquely characterised by the index function used and the data dimension, and thus do not vary across n\\_jellies and max\\_tries. Variables I\\_max\\_max, P\\_J\\_hat, and time are calculated at each 51 observations.
load(here::here("data/sim_df.rda"))
sim_df |> 
  dplyr::select(index, d, smoothness, squintability, n_jellies, max_tries, 
                I_max_max, P_J_hat, time) |> 
  arrange(index, d, n_jellies, max_tries) |> 
  head(5) |> 
  kable(digits = 3)
```

>>>>

```{r setup, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(ggh4x)
library(broom)
```

@@

[TODO: discuss what the values of smooth and theta3 mean for the optimisation]

@@

```{r}
#| label: tbl-smoothness-squintability
#| tbl-cap: Parameters estimated from the Gaussian process (including variance, range, smoothness, and nugget) and scaled logistic function (theta1 - theta4) for the pipe-finding and sine-wave finding problems. The squint column is calculated as theta1 * theta2 * theta3 / 4, as described in @sec-squintability. The "smooth" and "squint" column represent the smoothness and squintability measures. 
load(here::here("data/smoothness.rda"))
load(here::here("data/squintability.rda"))
library(kableExtra)
tibble(shape = c(rep("pipe", 4), rep("sine", 8)), smoothness) |>
  left_join(squintability) |> 
  rename(d = n, smooth = smoothness) |> 
  knitr::kable(digits = 3) |> 
  column_spec(1, border_left = TRUE) |> 
  column_spec(c(3, 7, 12), border_right = TRUE) |> 
  kable_styling(latex_options = "hold_position")
```

@@

```{r}
#| label: tbl-all-sim
#| tbl-cap: The first few rows of the datasets processed for modelling. The smoothness and squintability variable are uniquely characterised by the index function used and the data dimension, and thus do not vary across n\\_jellies and max\\_tries. The variable P\\_J\\_hat, and time are calculated at each observation.
load(here::here("data/sim_df.rda"))
sim_df |> 
  dplyr::select(index, d, smoothness, squintability, n_jellies, max_tries, 
                -I_max_max, P_J_hat, time) |> 
  arrange(index, d, n_jellies, max_tries) |> 
  head(7) |> 
  kable(digits = 3)
```

```{r}
sim_df2 <- sim_df |> 
  mutate(n_jellies = n_jellies/10,
         max_tries = max_tries/10,
         long_time = ifelse(time > 30, 1, 0),
         P_J_hat = ifelse(index == "stringy", 0, P_J_hat))

mod1 <- glm(P_J_hat ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
broom::tidy(mod1)
```

### 2024-06-03 optim.tex
finalise the calculation for example 2

\begin{tcolorbox}[enhanced jigsaw, left=2mm, coltitle=black, opacitybacktitle=0.6, colback=white, breakable, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, toptitle=1mm, bottomtitle=1mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomrule=.15mm, rightrule=.15mm, opacityback=0]

@@

@@

\begin{longtable}[t]{|>{}ll>{}r|rrr>{}r|rrr>{}r|}

\caption{\label{tbl-smoothness-squintability}Parameters estimated from
the Gaussian process (including variance, range, smoothness, and nugget)
and scaled logistic function (theta1 - theta4) for both the pipe-finding
and sine-wave finding problems. The ``smooth'' and ``theta3'' column
represent the smoothness and squintability measures.}

@@

\toprule
shape & index & d & variance & range & smooth & nugget & theta1 & theta2 & theta3 & theta4\\
\midrule
pipe & holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 & 1.001 & 0.860 & 3.368 & 0.823\\
pipe & holes & 8 & 0.000 & 0.259 & 2.373 & 0.613 & 1.001 & 0.869 & 3.264 & 0.811\\
pipe & holes & 10 & 0.000 & 0.144 & 2.317 & 1.831 & 1.000 & 0.885 & 3.151 & 0.806\\
pipe & holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 & 1.000 & 0.878 & 3.345 & 0.806\\
sine & MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 & 0.894 & 0.571 & 1.623 & -0.024\\
\addlinespace
sine & MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 & 0.932 & 0.328 & 1.314 & -0.030\\
sine & TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 & 0.951 & 0.536 & 1.719 & -0.025\\
sine & TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 & 0.945 & 0.564 & 1.723 & -0.027\\
sine & dcor2d\_2 & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019\\
sine & loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 & 1.016 & 1.039 & 2.648 & 0.080\\
\addlinespace
sine & splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 & 1.014 & 1.051 & 2.730 & -0.009\\
sine & stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 & 1.011 & 0.011 & 254.734 & 0.727\\
\bottomrule

@@

\begin{table}

\caption{\label{tbl-all-sim}The first five rows of the datasets after
combining all the calculation. The smoothness and squintability variable
are uniquely characterised by the index function used and the data
dimension, and thus do not vary across n\_jellies and max\_tries.
Variables I\_max\_max, P\_J\_hat, and time are calculated at each 51
observations.}

\end{table}%

>>>>

\begin{tcolorbox}[enhanced jigsaw, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, leftrule=.75mm, arc=.35mm, left=2mm, opacityback=0, breakable, toptitle=1mm, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, colback=white, titlerule=0mm, bottomrule=.15mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame]

@@

{[}TODO: discuss what the values of smooth and theta3 mean for the
optimisation{]}

@@

\begin{longtable}[t]{|>{}ll>{}r|rrr>{}r|rrrr>{}r|}

\caption{\label{tbl-smoothness-squintability}Parameters estimated from
the Gaussian process (including variance, range, smoothness, and nugget)
and scaled logistic function (theta1 - theta4) for the pipe-finding and
sine-wave finding problems. The squint column is calculated as theta1 *
theta2 * theta3 / 4, as described in Section~\ref{sec-squintability}.
The ``smooth'' and ``squint'' column represent the smoothness and
squintability measures.}

@@

\toprule
shape & index & d & variance & range & smooth & nugget & theta1 & theta2 & theta3 & theta4 & squint\\
\midrule
pipe & holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 & 1.001 & 0.860 & 3.368 & 0.823 & 0.725\\
pipe & holes & 8 & 0.000 & 0.259 & 2.373 & 0.613 & 1.001 & 0.869 & 3.264 & 0.811 & 0.710\\
pipe & holes & 10 & 0.000 & 0.144 & 2.317 & 1.831 & 1.000 & 0.885 & 3.151 & 0.806 & 0.697\\
pipe & holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 & 1.000 & 0.878 & 3.345 & 0.806 & 0.734\\
sine & MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 & 0.894 & 0.571 & 1.623 & -0.024 & 0.207\\
\addlinespace
sine & MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 & 0.932 & 0.328 & 1.314 & -0.030 & 0.100\\
sine & TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 & 0.951 & 0.536 & 1.719 & -0.025 & 0.219\\
sine & TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 & 0.945 & 0.564 & 1.723 & -0.027 & 0.230\\
sine & dcor2d\_2 & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019 & 0.679\\
sine & loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 & 1.016 & 1.039 & 2.648 & 0.080 & 0.699\\
\addlinespace
sine & splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 & 1.014 & 1.051 & 2.730 & -0.009 & 0.727\\
sine & stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 & 1.011 & 0.011 & 254.734 & 0.727 & 0.711\\
\bottomrule

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0811}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0405}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1486}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1892}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1351}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1351}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1081}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1622}}@{}}

\caption{\label{tbl-all-sim}The first few rows of the datasets processed
for modelling. The smoothness and squintability variable are uniquely
characterised by the index function used and the data dimension, and
thus do not vary across n\_jellies and max\_tries. The variable
P\_J\_hat, and time are calculated at each observation.}

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
index
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
d
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
smoothness
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
squintability
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
n\_jellies
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
max\_tries
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
P\_J\_hat
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
time
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
MIC & 6 & 2.394 & 0.207 & 20 & 50 & 0.12 & 2.479 secs \\
MIC & 6 & 2.394 & 0.207 & 20 & 100 & 0.24 & 8.950 secs \\
MIC & 6 & 2.394 & 0.207 & 50 & 50 & 0.52 & 5.651 secs \\
MIC & 6 & 2.394 & 0.207 & 50 & 100 & 0.64 & 13.223 secs \\
MIC & 6 & 2.394 & 0.207 & 100 & 50 & 0.76 & 19.453 secs \\
MIC & 8 & 2.394 & 0.100 & 20 & 50 & 0.08 & 2.566 secs \\
MIC & 8 & 2.394 & 0.100 & 20 & 100 & 0.08 & 4.960 secs \\

\begin{verbatim}
# A tibble: 7 x 5
  term          estimate std.error statistic p.value
  <chr>            <dbl>     <dbl>     <dbl>   <dbl>
1 (Intercept)     -4.67      4.51     -1.04  0.300  
2 smoothness       1.60      1.53      1.05  0.294  
3 squintability    7.06      2.21      3.20  0.00140
4 d               -0.595     0.260    -2.29  0.0221 
5 long_time       -0.851     1.35     -0.632 0.528  
6 n_jellies        0.230     0.131     1.75  0.0795 
7 max_tries        0.107     0.153     0.700 0.484  
\end{verbatim}
